---
title: "Beer Ingredients Plot"
author: "Tracy"
date: "January 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

This analysis is to do a replicative study of [shrin's playground](https://shiring.github.io/animation/2016/12/22/alcohol_post) on world's beer ingredients and alcohol and tabacco data from tax and trade bureau(TTB) (https://www.ttb.gov/).

The data comes in by month, year and following ingredients: 

* Malt and malt products
* Corn and corn products
* Rice and rice products
* Barley and barley products
* Wheat and wheat products
* Sugar and syrups
* Dry hops
* Hops extracts
* Other ingredients, like flavors, etc

## Read in data 

**First of al**l, let's read in the data and order it by year and calculate the percentage change from previous month. 


```{r, message=FALSE, warning=FALSE}
library(ttbbeer)
data("beermaterials")

library(dplyr)
# str(beermaterials)
# order by year from least to the most recent
beermaterials_change=arrange(beermaterials,Year)
# value=beermaterials_change[[colnames(beermaterials_change)[3]]] #that's how you get your value. by indexing your column names using df[[colname]]
# value[1]
# beermaterials_change[,3]#this will show the whole list value of that column 
for (i in 3:ncol(beermaterials_change)) {
  value=beermaterials_change[[colnames(beermaterials_change)[i]]]
  beermaterials_change[,i]=c(0,100*round((value[-1]-value[1:(length(value)-1)])/value[1:(length(value)-1)],2)) 
  
  #the above formula (value[-1]-value[1:(length(value)-1)])/value[1:(length(value)-1)] is trying to get the percentage of this month's increase rate from last month
  #this is to substitute the column value for the percentage
}

# head(beermaterials_change)

beermaterials$sum=rowSums(beermaterials[,-c(1:2)])
```


**Second**, let's change the data format from wide to long, in other words, simplify the data layout by just including 'ingredients' and 'value'.  


```{r}


library(tidyr)
# gather is great to change wide format to long format. 
beermaterials_gather=beermaterials %>%
  gather(Ingredients,Amount, Malt_and_malt_products:Other)
beermaterials_gather$Date=paste('01',beermaterials_gather$Month, beermaterials_gather$Year,sep = "-") %>%
  as.Date(format="%d-%B-%Y")
beermaterials_gather$Ingredients=gsub("_"," ",beermaterials_gather$Ingredients)

beermaterials_change_gather=beermaterials_change %>%
  gather(Ingredients,Amount, Malt_and_malt_products:Other)
beermaterials_change_gather$Date=paste('01',beermaterials_change_gather$Month, beermaterials_change_gather$Year,sep = "-") %>%
  as.Date(format="%d-%B-%Y")
beermaterials_change_gather$Ingredients=gsub("_"," ",beermaterials_change_gather$Ingredients)


```

## Exploratory analysis


Now let's do some exploratory analysis to see what the data looks like. Here we also set up the theme for the entire analysis. 

**First**,  we'd like to see the general trend of the beer ingredients by lumping them all together by 6 month interval. 

```{r}
# install.packages('ggplot2')
library(ggplot2)

my_theme=function(base_size=12,base_family="sans"){
  theme_minimal(base_size=base_size,base_family=base_family)+
    theme(
      axis.text=element_text(size=12),
      axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),
      axis.title=element_text(size=14),
      panel.grid.major=element_line(color='grey'),
      panel.grid.minor=element_blank(),
      panel.background=element_rect(fill="aliceblue"),
      strip.background=element_rect(fill="royalblue",color='grey',size=1),
      strip.text=element_text(face='bold',size=12,color='white'),
      legend.position='bottom',
      legend.justification='top',
      legend.box='horizontal',
      legend.box.background=element_rect(colour='grey50'),
      legend.background=element_blank(),
      panel.border=element_rect(color='grey',fill=NA,size=0.5)
      
  )
  
}

theme_set(my_theme())
# install.packages('lattice')
library(lattice)
ggplot(beermaterials_gather,aes(x=Date,y=sum))+
  geom_point(size=1.5,alpha=0.6,color='navy')+
  geom_smooth(alpha=0.3,color='black',size=0.5,linetype='dashed',method = 'gam')+
  geom_line(color='navy')+
  guides(color=FALSE)+
  labs(
    x="",
    y='Amount in pounds',
    title='Sum of Beer Ingredients by Month and Year',
    subtitle="Sum of monthly aggregates of all ingredients from 2006 to 2015"
  )+
  scale_x_date(date_breaks = '6 month',date_labels = "%B %Y")

```


**Second**, let's look at the amount of beer ingredients used over the time by 6 month interval: 

```{r}

ggplot(beermaterials_gather,aes(x=Date,y=Amount,colour=Ingredients))+
  geom_point(size=1.5,alpha=0.6)+
  geom_smooth(alpha=0.3,color='black',size=0.5,linetype='dashed')+
  geom_line()+
  guides(color=FALSE)+
  facet_wrap(~Ingredients,ncol = 3,scales = 'free_y')+
  labs(
    x="",
    y='Amount in pounds',
    title='Beer Ingredients by Month and Year',
    subtitle="monthly aggregates of all ingredients from 2006 to 2015"
  )+
  scale_x_date(date_breaks = '6 month',date_labels = "%B %Y") # date_breaks can specify any date interval by using quotes and text inside to indicate.


```


**thirdly**, let's zoom into more granular data by looking at it by 1 month interval and just display all the ingredients in one column:

```{r}

ggplot(beermaterials_gather,aes(x=Date,y=Amount,colour=Ingredients))+
  geom_point(size=1.5,alpha=0.6)+
  # geom_smooth(alpha=0.3,color='black',size=0.5,linetype='dashed')+
  geom_line()+
  guides(color=FALSE)+
  facet_wrap(~Ingredients,ncol = 1,scales = 'free_y')+
  labs(
    x="",
    y='Amount in pounds',
    title='Beer Ingredients by Month and Year',
    subtitle="monthly aggregates of all ingredients from 2006 to 2015"
  )+
  scale_x_date(date_breaks = '1 month',date_labels = "%B %Y") # date_breaks can specify any date interval by using quotes and text inside to indicate.



```

**fourthly**, let's look at the monthly percentage change by different ingredients:

```{r}

ggplot(beermaterials_change_gather,aes(x=Date,y=Amount,colour=Ingredients))+
  geom_point(size=1.5,alpha=0.6)+
  geom_smooth(alpha=0.3,color='black',size=0.5,linetype='dashed')+
  geom_line()+
  guides(color=FALSE)+
  facet_wrap(~Ingredients,ncol = 3,scales = 'free_y')+
  labs(
    x="",
    y='Percentage Change',
    title='Beer Ingredients by Month and Year',
    subtitle="Percentage Change of Monthly Aggregates of all ingredients from 2006 to 2015"
  )+
  scale_x_date(date_breaks = '1 year',date_labels = "%Y")



```

**next**, let's look at the percent change in pie format. 
```{r}
beeraterials_percent=cbind(beermaterials[,1:2],prop.table(as.matrix(beermaterials[,-c(1:2)]),margin=1)*100) #prop.table is used to express table as a fraction of marginal table. if margin=1, then the matrix row sum will become 1.
beermaterials_percent_gather=beeraterials_percent %>%
  gather(Ingredients,Amount, Malt_and_malt_products:Other)
beermaterials_percent_gather$Ingredients=gsub("_"," ",beermaterials_percent_gather$Ingredients)

#below is to make a customized factor for a certain variable
f=unique(beermaterials_percent_gather$Month)
beermaterials_percent_gather=within(beermaterials_percent_gather,{
  Month=factor(Month,levels = f)
  }
  )

```

Let's look at some animated charts:

```{r}
library(ggplot2)
library(animation)
ani.options(ani.width=800,ani.height=500)
saveGIF({
  for (year in rev(unique(beermaterials_percent_gather$Year))) {
    pie=ggplot(subset(beermaterials_percent_gather,Year==paste(year)),aes(x="",y=Amount,fill=Ingredients,frame=Year))+
      geom_bar(width=1,stat="identity")+theme_minimal()+coord_polar("y",start = 0)+
      labs(
        title="Percentage of Beer Ingredients by Year and Month",
        subtitle=paste(year))+
      #theme has to contain the below things but you can use element_blank() to mark empty
      theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.border = element_blank(),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(size = 14,face='bold'),
        legend.title = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 8))+
      facet_wrap(~Month,ncol = 6)+guides(fill=guide_legend(ncol = 9,byrow = F))
      print(pie)
  }
},movie.name = 'beer_ingredients.gif')

```
